---
title: "Initial Korean covid 19"
author: "Johana Quintanilla"
date: "10/14/2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When we plotted this information, we saw that the number of people alive is higher than the number of people dead. Additionally, when looking at the percentage chart, we found that in the chart of dead people, two thirds were men and one third women. This indicates that the disease killed males faster than females. In the alive chart, most women were still alive compared to men. 
  





## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Theses numerical results tell us that the percentage of females dying due to the COVID-19 disease is 1.4% while for the males it is 3%. This indicates that males have a higher chance of dying from COVID-19 than females do.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When we observed the inferential results, we found that the odds of males dying due to the COVID-19 disease in South Korea is 2.3 times higher than the females. This agrees with our past results since we saw that most of the people that died because of COVID are males. We also noticed that the P-value is low which means that the chances of males dying is higher than the chances of females dying due to the COVID-19 disease. 

## Conclusions

In this assignment, we were able to analyze COVID-19 data from South Korea. We noticed that this was some categorical-categorical data, so we used our helper file label Cat~Cat-3.rmd to help us answer some questions. We were able to analyze the data using graphical, numerical and inferential results. After analyzing our plots and tables, we concluded that males have a higher chance of dying from COVID-19 than females do. 
